TCP Congestion Algorithms In Datacenters \\ \vspace{0.3cm} {\large LS Project}

### 2017-12-11 paper/paper.tex
m. paper.tex: remove trailing spacing
m. paper.tex: unindent sub(sub)sections as discussed

\section{Method}
During this research a performance test of various TCP congestion algorithms will be performed. The most common TCP congestion algorithms will be determined through desk research. The tests will be conducted in a private environment. Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.

	\subsection{Selecting TCP congestion algorithms}	
	In order to select the most common TCP congestion algorithms desk research will be performed. This desk research will show which algorithms are used more often than other algoritms by looking for the default algorithms in Operating Systems. Based on those results a list of TCP congestion algorithms will be composed.

	\subsection{Comparing Performance}
	One scenario will be used to compare the performance of the TCP congestion algorithms:
	\begin{enumerate}
		\item The user owns multiple endpoints in the network. In this case a congestion algorithm is applied at both the client and server side. This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network. In this test the user node and random node(s) are at the receiving side and the user server and random server at the sending side. The tests will be performed from a multitenant datacenter perspective.
	\end{enumerate}

@@

		\subsubsection{Determining the Packet loss and Delay}
		The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research. The dataset includes global information and shows averages per month over the year 2017. The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country. Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:packet-loss-chart}

@@

		\subsubsection{Performing time-based automated tests}
		The tests will be performed using scripts written in Python [\ref{appendix:python}]. The scripts either act as controller or as disciple. The controller is responsible for controlling the tests, by synchronizing the time, exchanging commands to run, modify or stop the tests and gathering data. The disciple is the listener and executes all task commands the controller sends. The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer and measurement tools on the disciple and closing the connection with the controller.

		\subsubsection{Testing hardware}
		The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}. The switch shown in table \ref{table:spec3} has all router related functionalities, which might impact the network speed, turned off.

@@

	\subsection{Traffic Management}
	Desk research will be conducted to gather information related to traffic management. The goal is to find a mechanism that is capable of allocating bandwidth equally between customers of a datacenter. This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.

\section{Results}
Type some text here.

	\subsection{TCP Congestion Algorithms}
	The most common TCP congestion algorithms are CTCP, CUBIC and BIC. CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}. CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}. BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}. DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}. DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.

		\subsubsection{Algorithm Characteristics}
		The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}. Those parameters are estimates defined by the BBR algorithm. $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time. $RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queueing or processing delays. Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$. As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased. If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore. The $RTT$ can never be lower than $RT_{prop}$... explain more on delay perhaps?.

@@

		The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:
		\begin{enumerate}
			\item Simple Marking at the Switch. Based on the marking treshold $K$ a packet is marked with the CE flag. This only occurs when the queue occupancy is greater than K at arrival time. The marking treshold $K$ specifies the minimum value of queue occupancy.
			\item ECN-Echo at receiver. The receiver running the DCTCP algorithm ACKs every packet and sets the ECN-Echo flag only if the packet was marked with CE. The ECN-Echo flag is used to notify the sender of congestion in the network.
			\item Controller at the sender. The sender has an estimate of the fraction of packets that are marked with CE. This is based on the actual fraction of packets that were marked, the weight given to the new estimate compared to the past weight. This results in $\alpha$. $\alpha$ is used to indicate the congestion on a network with a maximum of value 1, indicating a congested network, and a minimum value of 0, indicating no congestion in the network. DCTCP cuts the window size using $\alpha$: $cwnd = cwnd * (1 - \alpha/2)$. This results in a low queue length, but ensuring high throughput.\\
		\end{enumerate}

		The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:
		\begin{enumerate}
			\item Binary search increase. This algorithm computes the midpoint between the maximum window size $W_{max}$ and current minimum window size $W_{min}$. If packet loss occurs, $W_{max}$ is set to the midpoint. This process repeats until $W_{max}$ and $W_{min}$ are smaller than the minimum increment $S_{min}$.
			\item Additive increase runs as an addition to the binary search increase. If $W_{max} - midpoint > S_{max}$, the window size is increased by $S_{max}$ until $W_{max} - midpoint < S_{max}$. $S_{max}$ is the preset maximum increment. 
		\end{enumerate}
In case the current window size becomes greater than $W_{max}$ a slow start strategy is performed to determine a new $W_{max}$: $W_{max} + \alpha * S_{min}$, where $\alpha$ is a multiplier. The multiplier is increased by one until $S_{min} * \alpha >= S_{max}$ or when losses occur. In the case of loss a new value for $W_{max}$ will be set.

@@

		\subsubsection{Performance comparison}	
		- Includes results of tests.

	\subsection{Traffic Management}
	- Includes what is available for trafic management in a datacenter.
	- Measures to take to equally allocate bandwidth to customers.
		\subsubsection{Quality of Service}

>>>>

\section{Method}

During this research a performance test of various TCP congestion algorithms will be performed. The most common TCP congestion algorithms will be determined through desk research. The tests will be conducted in a private environment. Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.

\subsection{Selecting TCP congestion algorithms}

In order to select the most common TCP congestion algorithms desk research will be performed. This desk research will show which algorithms are used more often than other algoritms by looking for the default algorithms in Operating Systems. Based on those results a list of TCP congestion algorithms will be composed.

\subsection{Comparing Performance}

One scenario will be used to compare the performance of the TCP congestion algorithms:

\begin{enumerate}
	\item The user owns multiple endpoints in the network. In this case a congestion algorithm is applied at both the client and server side. This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network. In this test the user node and random node(s) are at the receiving side and the user server and random server at the sending side. The tests will be performed from a multitenant datacenter perspective.
\end{enumerate}

@@

\subsubsection{Determining the Packet loss and Delay}

The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research. The dataset includes global information and shows averages per month over the year 2017. The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country. Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:packet-loss-chart}

@@

\subsubsection{Performing time-based automated tests}

The tests will be performed using scripts written in Python [\ref{appendix:python}]. The scripts either act as controller or as disciple. The controller is responsible for controlling the tests, by synchronizing the time, exchanging commands to run, modify or stop the tests and gathering data. The disciple is the listener and executes all task commands the controller sends. The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer and measurement tools on the disciple and closing the connection with the controller.

\subsubsection{Testing hardware}

The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}. The switch shown in table \ref{table:spec3} has all router related functionalities, which might impact the network speed, turned off.

@@

\subsection{Traffic Management}

Desk research will be conducted to gather information related to traffic management. The goal is to find a mechanism that is capable of allocating bandwidth equally between customers of a datacenter. This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.

\section{Results}

Type some text here.

\subsection{TCP Congestion Algorithms}

The most common TCP congestion algorithms are CTCP, CUBIC and BIC. CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}. CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}. BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}. DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}. DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.

\subsubsection{Theoretical Algorithm Characteristics}

The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}. Those parameters are estimates defined by the BBR algorithm. $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time. $RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queueing or processing delays. Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$. As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased. If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore. The $RTT$ can never be lower than $RT_{prop}$... explain more on delay perhaps?.

@@

The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:

\begin{enumerate}
	\item Simple Marking at the Switch. Based on the marking treshold $K$ a packet is marked with the CE flag. This only occurs when the queue occupancy is greater than K at arrival time. The marking treshold $K$ specifies the minimum value of queue occupancy.
	\item ECN-Echo at receiver. The receiver running the DCTCP algorithm ACKs every packet and sets the ECN-Echo flag only if the packet was marked with CE. The ECN-Echo flag is used to notify the sender of congestion in the network.
	\item Controller at the sender. The sender has an estimate of the fraction of packets that are marked with CE. This is based on the actual fraction of packets that were marked, the weight given to the new estimate compared to the past weight. This results in $\alpha$. $\alpha$ is used to indicate the congestion on a network with a maximum of value 1, indicating a congested network, and a minimum value of 0, indicating no congestion in the network. DCTCP cuts the window size using $\alpha$: $cwnd = cwnd * (1 - \alpha/2)$. This results in a low queue length, but ensuring high throughput.\\
\end{enumerate}

The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:

\begin{enumerate}
	\item Binary search increase. This algorithm computes the midpoint between the maximum window size $W_{max}$ and current minimum window size $W_{min}$. If packet loss occurs, $W_{max}$ is set to the midpoint. This process repeats until $W_{max}$ and $W_{min}$ are smaller than the minimum increment $S_{min}$.
	\item Additive increase runs as an addition to the binary search increase. If $W_{max} - midpoint > S_{max}$, the window size is increased by $S_{max}$ until $W_{max} - midpoint < S_{max}$. $S_{max}$ is the preset maximum increment.
\end{enumerate}
In case the current window size becomes greater than $W_{max}$ a slow start strategy is performed to determine a new $W_{max}$: $W_{max} + \alpha * S_{min}$, where $\alpha$ is a multiplier. The multiplier is increased by one until $S_{min} * \alpha >= S_{max}$ or when losses occur. In the case of loss a new value for $W_{max}$ will be set.

@@

\subsubsection{Performance comparison}

- Includes results of tests.

\subsection{Traffic Management}

- Includes what is available for trafic management in a datacenter.
- Measures to take to equally allocate bandwidth to customers.

\subsubsection{Quality of Service}
